Dispersion of short waves riding on longer waves

### 2024-06-07 paper/paper.tex
Updates on numerical results

TODO amplitude modulation;

@@

Long waves induce vertical orbital accelerations, so the short
waves that ride on their surface move in an accelerated reference frame and
experience effective gravitational acceleration that is lower at the crests and
higher in the troughs \cite{longuet1986eulerian,longuet1987propagation}.
A linearized (sine-like) wave has a modulation in gravity acceleration at its
surface:

\begin{equation}
\label{eq:effective_gravity}
\widetilde{g}(\psi) = g + \dfrac{\partial W}{\partial t} = g \left[
  1 - \varepsilon_L e^{\varepsilon_L \cos{\psi}} \left(
    \cos{\psi} - \varepsilon_L \sin^2{\psi}
  \right)
\right]
\end{equation}
The first-order term in $a_L k_L$ is in phase with the long-wave elevation,
so the apparent gravity of the short waves is increased at the long-wave crests
and decreased in the troughs.
The second-order term in $a_L k_L$ is positive-definite and phase-shifted by
$\pi/2$ relative to the crests.
The second-order term thus increases the apparent gravity on the front and rear
faces of the long wave (Fig. \ref{fig:apparent_gravity}).

@@

With the new long-wave modulated expressions for the short-wave wavenumber
(Eq. \ref{eq:k_short3}) and gravity (Eq. \ref{eq:apparent_gravity2}), we can
evaluate the long-wave modulated frequency as simply
$\widetilde{\omega}(\psi) = \sqrt{\widetilde{g} \widetilde{k}}$, which we compare
for four different values of $a_L k_L$ in Fig. \ref{fig:frequency_modulation}.

\begin{figure}[h]
\label{fig:frequency_modulation}
\centering
\includegraphics[width=\textwidth]{figures/fig_frequency_modulation_2panel.png}
\caption{
  Relative change in short-wave frequency due to the presence of longer
  waves, for four different values of long-wave steepness.
}
\end{figure}

\subsection{Frequency and group speed modulation}

@@

We now proceed to integrate the full numerical solutions for the short-wave
wavenumber and action density in presence of long waves.
Equations (\ref{eq:wavenumber})-(\ref{eq:wave_action2}) are discretized using
2nd order central finite difference in space and integrated in time using the
4th order Runge-Kutta method \cite{butcher1996runge}.
Although the 2nd order central finite difference is not appropriate for many
numerical problems, in our case it is sufficient because it is conservative and
the fields that we compute derivatives of are smooth and continuous.
The space is divided into 100 grid points.

@@

\begin{figure}[h]
\label{fig:frequency_modulation}
\centering
\includegraphics[width=\textwidth]{figures/fig_modulation_3panel_infinite_wave_train.png}
\caption{
  TODO
}
\end{figure}

\begin{figure}[h]
\label{fig:frequency_modulation}
\centering
\includegraphics[width=\textwidth]{figures/fig_modulation_3panel_linear_ramp.png}
\caption{
  TODO
}
\end{figure}

@@

\begin{figure}[h]
\label{fig:frequency_modulation}
\centering
\includegraphics[width=\textwidth]{figures/fig_modulation_3panel_wave_group.png}
\caption{
  TODO
}
\end{figure}

>>>>

TODO amplitude modulation;
The modulation of short-wave amplitude can be derived in the same way by
dropping the non-linear terms in the wave action balance equation
(Eq. \ref{eq:wave_action2}) and integrating wave action in time:

\begin{equation}
\label{eq:wave_action_modulation}
\widetilde{N}(\psi) = N e^{\varepsilon_L \cos{\psi} e^{\varepsilon_L \cos{\psi}}}
\end{equation}

Since wave action is variance divided by the intrinsic frequency $\sigma$,
and variance scales with the square of the amplitude, the modulation of
amplitude is proportional to the square root of the modulation of wave action.

\begin{equation}
\label{eq:wave_amplitude_modulation}
\widetilde{a}(\psi) = a \sqrt{\dfrac{\sigma}{\widetilde{\sigma}} \dfrac{\widetilde{N}}{N}}
\end{equation}
To determine $\widetilde{\sigma}$, we will need to evaluate the modulation of
the gravitational acceleration at the surface of the long wave.

@@

Long waves induce vertical orbital accelerations, so the short
waves that ride on their surface move in an accelerated reference frame and
experience effective gravitational acceleration that is lower at the crests and
higher in the troughs \cite{longuet1986eulerian,longuet1987propagation}.
A linear wave causes a modulation in gravity acceleration at its surface:

\begin{equation}
\label{eq:gravity_modulation_linear}
\widetilde{g}(\psi) = g + \dfrac{\partial W}{\partial t} = g \left[
  1 - \varepsilon_L e^{\varepsilon_L \cos{\psi}} \left(
    \cos{\psi} - \varepsilon_L \sin^2{\psi}
  \right)
\right]
\end{equation}
The first-order term in $a_L k_L$ is in phase with the long-wave elevation,
so the apparent gravity of the short waves is increased at the long-wave crests
and decreased in the troughs.
The second-order term in $a_L k_L$ is positive-definite and phase-shifted by
$\pi/2$ relative to the crests.
The second-order term thus increases the apparent gravity on the front and rear
faces of the long wave (Fig. \ref{fig:gravity_modulation}).

@@

We can also derive the short-wave gravitational acceleration on the surface of
a non-linear Stokes wave with the elevation (to the third order in $\varepsilon_L$):

\begin{equation}
\label{eq:eta_stokes}
\eta_{St} = a_L \left[
  \cos{\psi} +
  \dfrac{1}{2} \varepsilon_L \cos{2\psi} +
  \varepsilon_L^2 \left( \dfrac{3}{8} \cos{3\psi} - \dfrac{1}{16} \cos{\psi} \right)
\right]
\end{equation}

The corresponding gravitational acceleration at the surface of the Stokes wave
is:

\begin{equation}
\label{eq:gravity_modulation_stokes}
\widetilde{g}(\psi) =
g \left\{
  1 - \varepsilon_L e^{k \eta_{St}}
  \left[ \cos{\psi} -
    \varepsilon_L \sin{\psi} \left(
      \sin{\psi}
      + \varepsilon_L \sin{2\psi}
      - \dfrac{1}{16} \varepsilon_L^2 \sin{\psi}
      + \dfrac{9}{8} \varepsilon_L^2 \sin{3\psi}
    \right)
  \right]
\right\}
\end{equation}

With the new long-wave modulated expressions for the short-wave wavenumber
(Eq. \ref{eq:k_short_exact}) and gravity
(Eqs. \ref{eq:gravity_modulation_linear}-\ref{eq:gravity_modulation_stokes}),
we can evaluate the long-wave modulated frequency as
$\widetilde{\omega}(\psi) = \sqrt{\widetilde{g} \widetilde{k}}$.
We compare the modulated intrinsic frequency for short waves on the surface of
linear and Stokes waves with $\varepsilon_L=2$ as function of phase, and their
phase-integral statistics in Fig. \ref{fig:frequency_modulation}.

\begin{figure}[h]
\label{fig:frequency_modulation}
\centering
\includegraphics[width=\textwidth]{figures/fig_frequency_modulation_2panel.png}
\caption{
  (a) Relative change in the short-wave intrinsic frequency on the surface of a
  longer linear (black) and Stokes (red) wave of steepness $\varepsilon_L = 0.2$
  as function of long-wave phase;
  (b) Long-wave phase maximum (solid), mean (dashed), and minimum (dotted)
  relative change in the short-wave intrinsic frequency as function of long-wave
  steepness in the case of a linear (black) and Stokes (red) long wave.
}
\end{figure}

@@

We now proceed to integrate the full numerical solutions for the short-wave
wavenumber and action density in presence of long waves, and in absence of
sources or sinks of wave energy.
Equations (\ref{eq:wavenumber})-(\ref{eq:wave_action2}) are discretized using
2$^{nd}$  order central finite difference in space and integrated in time using the
4$^{th}$ order Runge-Kutta method \cite{butcher1996runge}.
The space is divided into 100 grid points.
This is effectively the same equation set and numerical configuration as that of
\citeA{peureux2021unsteady}, except that, like \citeA{longuet1960changes}, their
long-wave orbital velocities are evaluated at $z=0$ instead of $z=\eta_L$.
The key difference is that evaluating orbital velocities at $z=0$ neglects the
Stokes drift of short-wave groups.
Although not significantly consequential for the modulation results here, there
is no good reason to exclude it given that its implementation is trivial.

Another difference is the choice of the numerical scheme for spatial differences,
which in their case is the MUSCL4 scheme by Kurganov and Tadmor (2000 TODO).
Although the 2$^{nd}$ order central finite difference is not appropriate for many
numerical problems, in our case it is sufficient because it is conservative and
the fields that we compute derivatives of are smooth and continuous.
We show in the next section that our numerical results in infinite long-wave
trains are qualitatively equivalent to those of \citeA{peureux2021unsteady}.

Finally, the numerical equations here are integrated in a fixed reference frame,
rather than one that moves with the long wave phase speed.
Either approach produces qualitatively equivalent results, but we find that
a fixed reference frame allows for more intuitive interpretation of the results.
However, this difference is important to keep in mind when comparing the results
of \citeA{peureux2021unsteady} and those presented here--in their Fig. 1 the
long-wave phase is fixed and the short waves are moving leftward
with the speed of $C_{pL} - C_g - U$ (neglecting any group speed inhomogeneity),
whereas in the figures that follow, the long wave is moving rightward with its
phase speed $C_{pL}$ and the short waves (that is, their energy) are moving
rightward with the speed of $C_g + U$ (neglecting any group wave inhomogeneity).

@@

The key result of \citeA{peureux2021unsteady} is that of unsteady steepening of
short waves when a homogeneous field of short waves is forced by an infinite
long-wave train.
The steepening is driven by the progressive increase in the action of short
waves (as opposed to their wavenumber), and is centered on the group of short
waves that begins its journey in the trough of the long wave.
This is because the unmodulated short waves that begin in the trough enter into
the area where the surface velocities are converging.
Inversely, the short waves that begin around the crest of the long-wave will
enter into the area of surface divergence and will thus have their wave action
and wavenumber, decreased.
We reproduce this behavior in Fig. \ref{fig:modulation_3panel_infinite}, which
shows the evolution of the short-wave action, wavenumber, and steepness
modulation (here defined as the change of a quantity relative to its initial value).
After 10 long-wave periods, the short-wave action modulation $\widetilde{N}/N$
is approximately doubled, which almost exactly corresponds to the unsteady growth
reported by \citeA{peureux2021unsteady} in their Fig. 1.
Another important property of this solution is that the short-wave group that
experiences peak amplification (dampening) are not locked-in to the long-wave
crests (troughs), as they are in the L-HS solution.

\begin{figure}[h]
\label{fig:modulation_3panel_infinite}
\centering
\includegraphics[width=\textwidth]{figures/fig_modulation_3panel_infinite_wave_train.png}
\caption{
  Change in the short-wave action (left), wavenumber (middle), and steepness (right)
  relative to their initial values as function of initial long-wave phase and time.
  $k_L = 1$, $k = 10$, $\varepsilon_L = 0.1$.
}
\end{figure}

Why does this unsteady growth of short-wave action occur?
As \citeA{peureux2021unsteady} explain, "the additional change in advection
velocity due to the modulation of the short wave group speed gives an additional
amplification."
That is, it is the inhomogeneity term $N \dfrac{\partial C_g}{\partial x}$ that
is responsible for the instability.
However, the instability only occurs if the short waves are initialized as a
uniform field.
If they are initialized instead as a periodic function of the long-wave phase
such that their wavenumber (and optionally, action) is higher on the crests
and lower in the troughs, akin to the L-HS solution, the instability does not
occur.

\begin{figure}[h]
\label{fig:modulation_3panel_ramp}
\centering
\includegraphics[width=\textwidth]{figures/fig_modulation_3panel_linear_ramp.png}
\caption{
  Same as Fig. \ref{fig:modulation_3panel_infinite} but with a linear ramp
  applied to $a_L$ during the initial five long-wave periods.
}
\end{figure}

Why is it, then, that the solution is so different depending on the short-wave
initial conditions?
\citeA{peureux2021unsteady} did not provide an explanation for this behavior.
At first, it may seem that there is something special about the uniform initial
conditions that makes its solutions grow unsteadily and indefinitely.
\citeA{peureux2021unsteady} describe this scenario as "...the sudden appearance
of a long wave perturbation in the middle of a homogeneous short wave field".
The sudden appearance of long waves may be a clue.
The exact numerical simulation as shown in Fig. \ref{fig:modulation_3panel_infinite},
except for a linear ramp applied to $a_L$ (and consequently, $U$ and $W$ as well)
for the first five periods reveals that the unsteady growh no longer occurs
(Fig. \ref{fig:modulation_3panel_ramp}).

\begin{figure}[h]
  \label{fig:modulation_3panel_ramp}
  \centering
  \includegraphics[width=\textwidth]{figures/fig_modulation_ramp_timeseries.png}
  \caption{
    Maximum short-wave steepness modulation as function of time in case of
    infinite long-wave train case (blue) and the same case but with a linear
    ramp during the initial five long-wave periods (orange). The dashed black
    line corresponds to the short-wave steepness of 0.4 at which most waves are
    expected to break.
  }
  \end{figure}

@@

\begin{figure}[h]
\label{fig:modulation_3panel_groups}
\centering
\includegraphics[width=\textwidth]{figures/fig_modulation_3panel_wave_group.png}
\caption{
  Same as Fig. \ref{fig:modulation_3panel_infinite} but for a long-wave group,
  causing the long-wave amplitude to gradually increase and peak at $a_L = 0.1$
  after five long-wave periods, and then conversely decay back to a calm sea state.
}
\end{figure}
